---
title: "Part 1 - MLR"
author: "Darreion Bailey"
date: "2024-04-28"
output: html_document
---

```{r setup, include=FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
library(corrplot)
library(tidyverse)
library(scales)
library(MASS)
library(leaps)
Data<-read.csv("kc_house_data.csv", sep=",", header=TRUE)
str(Data)
summary(Data)
colSums(is.na(Data))
```

```{r}

Data$sqft_bathrooms<-Data$sqft_living * Data$bathrooms #interaction term

set.seed(6021)
sample.data<-sample.int(nrow(Data), floor(.50*nrow(Data)), replace = F)
train<-Data[sample.data, ]
#Add a column to indicate either house is renovated or constructed after 1999 or not. 
train$construction_category <- ifelse(train$yr_built >= 2000 | train$yr_renovated >= 2000, 1, 0)

test<-Data[-sample.data, ]
#Add a column to indicate either house is renovated or constructed after 1999 or not. 
test$construction_category <- ifelse(test$yr_built >= 2000 | test$yr_renovated >= 2000, 1, 0)
train<-dplyr::select(train, -id, -sqft_above, -sqft_basement, -date, -yr_renovated, -yr_built, -lat, -long, -zipcode)
# removing variables not associated with home features
matrix<-corrplot(cor(train),
         method = "number",
         type = "upper",
         number.cex=0.50)

```
```{r}

train<-dplyr::select(train, -sqft_lot15,)

```
Also remove sqft_lot15 since sqft_living15 likely covers similar information, and we want to capture urban effects which living square feet might be better for.

```{r}
full_model <- lm(price~., data=train)
backward_model <- step(full_model, direction = "backward")
forward_model <- step(full_model, direction = "forward")
stepwise_model <- step(full_model, direction = "both")
AIC(backward_model)
AIC(forward_model)
AIC(stepwise_model)
```
Move forward with stepwise model because tied with forward for lowest AIC.

```{r}
#Residuals vs Fitted
plot(stepwise_model, which=1)
```

```{r}
## QQ Plot
plot(stepwise_model,which=2)
```

```{r}
plot(stepwise_model, which = 5)
```

Given the potential non-linearity shown in the residuals vs fitted plot, we will try to log transform the price variable to see if that helps the fit. Increasing variance in residual plot indicates we should try a transformation on the response variable. Box cox plot shows lambda very near to zero, so we use a log transformation on price. 


```{r}
MASS::boxcox(stepwise_model)
train$price_log<-log(train$price)
model_log_trans<-lm(price_log~. -price, data=train)
stepwise_model_trans<- step(model_log_trans, direction = "both")
summary(stepwise_model_trans)
```


Now let's run the plots again

```{r}
#Residuals vs Fitted
plot(model_log_trans, which=1)
```

```{r}
plot(model_log_trans,which=2)
```
```{r}
plot(model_log_trans, which = 5)
```

```{r}
specific_point1 <- Data[15871, ]
specific_point1

specific_point2<-Data[12778, ]
specific_point2

specific_point3<-Data[9715, ]
specific_point3

specific_point4<-Data[1720, ]
specific_point4
```
Here we identify high leverage observations and outliers. The data point 15871 must be an entry error, because with the sqft_living being 1620sqft, it is not possible for there to be 33 bedrooms. Therefore, we can remove this entry
```{r}
hii<-lm.influence(stepwise_model)$hat ##leverages
ext.student<-rstudent(stepwise_model) ##ext studentized res
n<-nrow(train)
p<-14

#Alot of obs
length(hii[hii>2*p/n])

length(ext.student[abs(ext.student)>3])

train<-train[train$bedrooms != 33, ]
```



Re-run regression

```{r}
train$price_log<-log(train$price)
model_log_trans<-lm(price_log~. -price, data=train)
stepwise_model_trans<- step(model_log_trans, direction = "both")
summary(stepwise_model_trans)
```
Still have a lot of high leverage obs, and 24 outliers. 
```{r}
hii2<-lm.influence(stepwise_model_trans)$hat ##leverages
ext.student2<-rstudent(stepwise_model_trans) ##ext studentized res
n<-nrow(train)
p2<-12

#Alot of obs
length(hii2[hii2>2*p2/n])

length(ext.student2[abs(ext.student2)>3])

```

```{r}
round(faraway::vif(stepwise_model_trans),3)
```
We observe sqft_living, sqft_bathrooms with VIFS over 10. Try removing sqft_bathrooms then refit. 

```{r}
model_log_trans<-lm(price_log~. -price -sqft_bathrooms, data=train)
stepwise_model_trans<- step(model_log_trans, direction = "both")
summary(stepwise_model_trans)
```
VIFS looks reasonable after removing sqft_bathrooms.
```{r}
round(faraway::vif(stepwise_model_trans),3)
```

```{r}
plot(stepwise_model_trans, which=1)
```

```{r}
plot(stepwise_model_trans,which=2)
```

```{r}
plot(stepwise_model_trans, which = 5)
```

```{r}
summary(stepwise_model_trans)
```
From these diagnostic plots, we can reasonably conclude the assumptions are met



Now we run vs. the test set

```{r}
test$predicted_log_price<-predict(stepwise_model_trans, newdata=test)
test$predicted_price <- exp(test$predicted_log_price)
rmse <- sqrt(mean((exp(test$predicted_log_price) - test$price)^2))
mae <- mean(abs(exp(test$predicted_log_price) - test$price))

plot(test$price, test$predicted_price, xlab = "Actual Price", ylab = "Predicted Price")
abline(0, 1)
```

The model is very good at predicting prices below $300,000, but as the outliers are introduced it has a harder time predicting those. The variance is rather high, and the RMSE and MAE are high as well for values greater than $300,000. Figure out how to simplify model further, and account for variance and outliers

